Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-07 thesis/paper.tex
Spellcheck

\author{
% 1st. author
\alignauthor
John Burchell\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{john.a.burchell@gmail.com}
% 2nd. author
\alignauthor
William Granli\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{william.granli@gmail.com}
\and
% 3rd. author
\alignauthor Imed Hammouda\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
% 4th 
\alignauthor Er ic Knauss\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
}

@@

%Structure of the paper
The sections following the Introduction are structured as follows: In Section \ref{existing_literature}, we present the background and previously published work that is related to our study. Section \ref{case_company_description} describes the case company and the API which was studied. A description of our methodology is introduced in Section \ref{methodology}. In Section \ref{results}, the results of our study are presented and in Section \ref{discussion}, we discuss the findings of our study. In Secton \ref{conclusion} we suggest the potential impact of our work and advise future work.

@@

\subsection{API Design} \label{api_design}
API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating} \cite{bloch2008effective} \cite{stylos2006comparing}. Examples of such decisions include whether or not to make use of the factory pattern or if an API should display errors only at compilation or only at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design decisions, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API is clear to the user, for example, using design patterns \cite{stylos2006comparing} \cite{shi2011empirical}. The design decisions reached during development of an API will effect the overall usability of the API, for better or worse. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and the API's user \cite{clarke2004measuring}.

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field that studies the application of software maintenance activities, changes in software processes and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s when it was first introduced to the software development community \cite{lientz1980software}. A set of four categories describing software maintenance categories \cite{lientz1980software} became the basis upon which twelve new categories were created \cite{chapin2001types}. The tweleve categories are: enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a software maintenance activity that relates to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. It has been suggested that software maintenance is a part of software evolution and that the two terms are not interchangeable \cite{chapin2001types}.

%%Lehman's laws (!!!improve later)
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman1980programs}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarities, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System \cite{lehman1980programs}. The laws are said to only apply to E-type systems, which are programs that are ``change prone" and that ``mechanise a human or societal activity" \cite{lehman1980programs}.

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study meth-odology \cite{runeson2009guidelines}. The goal of the study was twofold. Firstly, the aim was to empirically \textit{explore} the motives behind changes to APIs and secondly, our ambition was to \textit{validate} the results' level of correctness by comparing our findings with the case company's explanation of their motives. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis.

@@

\subsubsection{axevent} %Reviewed by JW
Version 2.0 of the API was designed to be an event-driven system that makes use of callbacks. To achieve this, the old event system had to be re-designed. The new design is used by all features of version 2.0 of the API. An example of when an event would be generated, is when the camera detects movement.  Given that this module was a re-implementation of the previous event system, we concluded that it should be categorised mainly as an enhancive change with the secondary change as groomative. Groomative was chosen as the secondary category due to the additional motive of bringing the new functionality into line with the new design. The interviewee outlined that the previous version of axevent, event, was deprecated in favour of a re-implemention of the new event system. One reason was that the case company was dissatisfied with certain usability aspects of API version 1.4. The module was then completely re-implemented and re-designed according to the new system design of version 2.0. From the discussion in the interview, it was clear that both of our categorisations were correct. The change was primarily enhancive and secondarily groomative.

@@

\subsubsection{3\textsuperscript{rd} Party}
%% REMEMBER TO PROVIDE LINK TO GLIB:::: - https://developer.gnome.org/glib/
ATTN: (!!!We are waiting for more information from Johanna to confirm / add to this section!!!)
%%Describe the change 
2.0 utilizes more 3rd party libraries
axsound, for example, uses alsa
All modules use glib
The advent of the new API saw the inclusion of new 3\textsuperscript{rd} party libraries. For example, axsound utilises a library called Alsa, all interfaces use Glib and daemons have been provided for using D-Bus. These changes potentially make the library much more flexible both for the API designers in terms of features they can provide and for the application developers in terms of applications they can develop.
We therefore gave this change the main categorisation of adaptive with no suitable sub-categories being found.
The architects creating the API decided they wished to use Glib which came about primarily from internal discussions at the company.
%%Relate it to Categories
Architects decided to use glib, part of an internal decision. 
Glib (!!!ref to website) was therefore introduced to the new API to help alleviate this 
The new API utilises Glib in the majority of its interfaces and considering the fact that Glib has its own type of error handling, it seemed sensible to implement the same technique into the new API. However, it was expressed to us that the error handling would have changed between versions regardless of whether or not Glib was included. (!!!This is also too assume-y)

@@

%%The following should probably be in conclusion, right? /Willy
       What this means for industry:
              Greater emphasis should be placed upon communicating with the end users / application developers. APIs should be made with maintainability and extendibility in mind as they seem to be what changes the most between versions. They should probably improve their feedback system to get more feedback from the API users.

@@

Considering the Ecosystem and considering it as a whole when designing the API. She does not think that they did take it into consideration at all. They had not thought about them when designing the APIs. She does think about who is going to use it and how but nothing in regards to how it affects an ecosystem as a whole. Things like allowing embedded webpages for the ACAP allows users to get more information from their system meaning they don’t need to ask for as much help and assistance. They usually integrate this with the VMS. Overall there was not much thought about ecosystem when the API was designed. She’s unsure if it’s one ecosystem or two. The network APIs and their API. Perhaps it is one ecosystem. (this is something Imed is looking at).
It seems that a focus on ecosystems might be starting to come through but that they already thought about them without calling them ecosystems. For example, her mentioning that they already did think about the users of the API and how they would be using it gives a slight clue that they do already think about the wider audience of their API.

@@

%%William add conclusion ideas here
It's interesting that the partners have mentioned that they want more frequent updates... maybe API developers should be less afraid of updating them? 
That the part about controlling how the API users use the API was such an important thing is interesting. (Not something we expected). 
-Relate to 80% refactorings paper and say that we found different results
*This is an intersting note if we want to compare platform apis and library apis to say that functionality changes more often in platform apis for example

@@

\section{Acknowledgments}
We would like to thank Axis and that project at GU.

>>>>

\author{
% 1st. author
\alignauthor
John Burchell\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{john.a.burchell@gmail.com}
% 2nd. author
\alignauthor
William Granli\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{william.granli@gmail.com}
\and
% 3rd. author
\alignauthor Imed Hammouda\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
% 4th 
\alignauthor Eric Knauss\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
}

@@

%Structure of the paper
The sections following the Introduction are structured as follows: In Section \ref{existing_literature}, we present the background and previously published work that is related to our study. Section \ref{case_company_description} describes the case company and the API which was studied. A description of our methodology is introduced in Section \ref{methodology}. In Section \ref{results}, the results of our study are presented and in Section \ref{discussion}, we discuss the findings of our study. In Section \ref{conclusion} we suggest the potential impact of our work and advise future work.

@@

\subsection{API Design} \label{api_design}
API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating} \cite{bloch2008effective} \cite{stylos2006comparing}. Examples of such decisions include whether or not to make use of the factory pattern or if an API should display errors only at compilation or only at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design decisions, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API is clear to the user, for example, using design patterns \cite{stylos2006comparing} \cite{shi2011empirical}. The design decisions reached during development of an API will effect the overall usability of the API, for better or worse. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and the user of it \cite{clarke2004measuring}.

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field that studies the application of software maintenance activities, changes in software processes and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s when it was first introduced to the software development community \cite{lientz1980software}. A set of four categories describing software maintenance categories \cite{lientz1980software} became the basis upon which twelve new categories were created \cite{chapin2001types}. The twelve categories are: enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a software maintenance activity that relates to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. It has been suggested that software maintenance is a part of software evolution and that the two terms are not interchangeable \cite{chapin2001types}.

%%Lehman's laws (!!!improve later)
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman1980programs}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarity, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System \cite{lehman1980programs}. The laws are said to only apply to E-type systems, which are programs that are ``change prone" and that ``mechanise a human or societal activity" \cite{lehman1980programs}.

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. The goal of the study was twofold. Firstly, the aim was to empirically \textit{explore} the motives behind changes to APIs and secondly, our ambition was to \textit{validate} the results' level of correctness by comparing our findings with the case company's explanation of their motives. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis.

@@

\subsubsection{axevent} %Reviewed by JW
Version 2.0 of the API was designed to be an event-driven system that makes use of callbacks. To achieve this, the old event system had to be re-designed. The new design is used by all features of version 2.0 of the API. An example of when an event would be generated, is when the camera detects movement.  Given that this module was a re-implementation of the previous event system, we concluded that it should be categorised mainly as an enhancive change with the secondary change as groomative. Groomative was chosen as the secondary category due to the additional motive of bringing the new functionality into line with the new design. The interviewee outlined that the previous version of axevent, event, was deprecated in favour of a re-implementation of the new event system. One reason was that the case company was dissatisfied with certain usability aspects of API version 1.4. The module was then completely re-implemented and re-designed according to the new system design of version 2.0. From the discussion in the interview, it was clear that both of our categorisations were correct. The change was primarily enhancive and secondarily groomative.

@@

\subsubsection{3\textsuperscript{rd} Party}
%% REMEMBER TO PROVIDE LINK TO GLIB:::: - https://developer.gnome.org/glib/
ATTN: (!!!We are waiting for more information from Johanna to confirm / add to this section!!!)
%%Describe the change 
2.0 utilises more 3rd party libraries
axsound, for example, uses alsa
All modules use glib
The advent of the new API saw the inclusion of new 3\textsuperscript{rd} party libraries. For example, axsound utilises a library called Alsa, all interfaces use Glib and daemons have been provided for using D-Bus. These changes potentially make the library much more flexible both for the API designers in terms of features they can provide and for the application developers in terms of applications they can develop.
We therefore gave this change the main categorisation of adaptive with no suitable sub-categories being found.
The architects creating the API decided they wished to use Glib which came about primarily from internal discussions at the company.
%%Relate it to Categories
Architects decided to use glib, part of an internal decision. 
Glib (!!!ref to website) was therefore introduced to the new API to help alleviate this 
The new API utilises Glib in the majority of its interfaces and considering the fact that Glib has its own type of error handling, it seemed sensible to implement the same technique into the new API. However, it was expressed to us that the error handling would have changed between versions regardless of whether or not Glib was included. (!!!This is also too assume-y)

@@

%%The following should probably be in conclusion, right? /Willy
       What this means for industry:
              Greater emphasis should be placed upon communicating with the end users / application developers. APIs should be made with maintainability and extensibility in mind as they seem to be what changes the most between versions. They should probably improve their feedback system to get more feedback from the API users.

@@

Considering the Ecosystem and considering it as a whole when designing the API. She does not think that they did take it into consideration at all. They had not thought about them when designing the APIs. She does think about who is going to use it and how but nothing in regards to how it affects an ecosystem as a whole. Things like allowing embedded web pages for the ACAP allows users to get more information from their system meaning they don’t need to ask for as much help and assistance. They usually integrate this with the VMS. Overall there was not much thought about ecosystem when the API was designed. She’s unsure if it’s one ecosystem or two. The network APIs and their API. Perhaps it is one ecosystem. (this is something Imed is looking at).
It seems that a focus on ecosystems might be starting to come through but that they already thought about them without calling them ecosystems. For example, her mentioning that they already did think about the users of the API and how they would be using it gives a slight clue that they do already think about the wider audience of their API.

@@

%%William add conclusion ideas here
It's interesting that the partners have mentioned that they want more frequent updates... maybe API developers should be less afraid of updating them? 
That the part about controlling how the API users use the API was such an important thing is interesting. (Not something we expected). 
-Relate to 80% refactorings paper and say that we found different results
*This is an interesting note if we want to compare platform APIs and library APIs to say that functionality changes more often in platform APIs for example

@@

\section{Acknowledgements}
We would like to thank Axis and that project at GU.
